An Adolescent Mini-Neptune in the Kepler Field

### 2021-07-17 paper/ms.tex
update ms

\documentclass[12pt,modern,tighten]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{---}
\revised{---}
\accepted{---}
\submitjournal{Nature}
\shorttitle{Kepler\,1627}

@@

\begin{abstract}
  Kepler\,1627A is a G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  selected for observation by the Kepler satellite because it is
  nearby ($d\approx 333\,{\rm pc}$) and it resembles the Sun.  Combining
  data from Gaia with rotation periods from TESS, we show that
  Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
  cluster.  Kepler\,1627Ab is therefore the youngest planet with a
  precise age found to date by the main Kepler mission.
  Analyzing the photometry, we find evidence that starspot crossings
  distort the average transit shape, and induce transit timing
  variations that suggest the planet is on a prograde orbit.  The 
  flare arrival time distribution is most likely Poissonian, though
  there may be hints ($<3$-$\sigma$) of additional flares at the
  orbital and synodic periods.  In the future, spectroscopic studies
  of \pn\ may help clarify the evolution of this mini-Neptune's orbit
  and atmosphere.  More broadly, the combination of Gaia and TESS data
  offers an opportunity to expand the census of age-dated planets; we
  provide a literature concatenation of young stars to support this
  effort.
\end{abstract}

\keywords{
  planetary evolution (XXXX),
  stellar associations (1582),
  open star clusters (1160),
	stellar ages (1581),
}

@@

\section{Manuscript}

While thousands of exoplanets have been discovered orbiting nearby
stars, the vast majority of them are several billion years old.  This
makes it difficult to test theories for the origins of the different
families of planets, since many evolutionary processes are expected to
operate on timescales of less than 100 million years.

For instance, the ``mini-Neptune'' planets, thought to be made of
molten rocky cores and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years
\citep{owen_constraining_2020,kite_atmosphere_2020}.  Specifically,
they start with sizes of 4--12\,$R_\oplus$ at the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which can eventually deplete or entirely strip the envelope
from the rocky core \citep{Owen_Wu_2013,gupta_sculpting_2019}.
Discovering young planets and measuring their compositions and
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of 
mature exoplanets \citep{Fulton_et_al_2017}.

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
    points are kinematically selected members of the $\delta$\,Lyr
    cluster.  Both Kepler (gray squares) and TESS (colored points)
    observed portions of the cluster.  Naked-eye stars ($m_{\rm
    V}<6.5$) are shown in gray; four of them (orange points) have
    their names annotated.  Kepler\,1627 (green star) was observed
    during the entirety of the Kepler mission, and has been observed
    by TESS for two lunar months to date.
		\label{fig:skychart}
	}
\end{figure*}

@@

@@

At the time of the main Kepler mission (2009--2013), only four open
clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Since that time, analyses of the
kinematic, photometric, and astrometric Gaia data have expanded our
knowledge of open cluster and moving group memberships \citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,kerr_stars_2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis by J.
Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
the cluster from the Kepler-52 and Kepler-968 systems
\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
discovery, and the focus of this work, is that Kepler observed a
portion of the $\delta$\,Lyr cluster \citep{stephenson_possible_1959,eggen_photometric_1968}.
The overlap between the Kepler field and the cluster is shown in
Figure~\ref{fig:skychart}.

{\bf Jason Curtis todo}
The \cn\ contains
Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470), which 
was previously reported to host a close-in mini-Neptune candidate planet
\citep{2012ApJS..199...24T,thompson_planetary_2018}
Given the planet's previous statistical validation by
\citet{morton_false_2016}, we decided to examine the age of the star
and its cluster more closely.  We collected the previously measured
{\bf XXX} Kepler rotation periods (CITEP MCQUILLAN2014), and measured
{\bf YYY} new stellar rotation periods for cluster members observed by
TESS.  We also performed an isochronal analysis.  The results are
shown in Figure~\ref{fig:age}.  From the ensemble perspective, which
provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
of the star is $XX \pm YY$\,Myr.  Individual measurements of the
star's rotation period and lithium abundance corroborate this
conclusion.

@@

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f11.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Evidence for a prograde orbit from transit timing variations.}
    The time of each transit was measured, along with the local slope
    of the light curve.  The two quantities appear anti-correlated,
    which is most easily explained by starspot crossings during the
    first (second) half of transit inducing a positive (negative) TTV,
    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
    units along the abscissa are most easily understood by considering
    that the flux from the star changes by $\sim$60\,ppt per half
    rotation period ($\sim$1.3\,days).
		\label{fig:ttvslope}
	}
\end{figure*}

@@

In the context of the known transiting planets, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precisely measured ages spanning 10 to 100\,Myr include K2-33b,
DS~Tuc~Ab, HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet
V1298~Tau system
\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020}.
Of these systems, \pn\ inhabits a similar range of planetary size and
orbital period.  It is the smallest though ($3.6\,R_\oplus$), which
may hint at selection effects in the sample of young planets
\citep{zhou_2021_tois}.  Assuming that these planets have masses
between $\approx$10 and 20$\,M_\oplus$, their large sizes and presumed
low densities are in accord with the expectation that mini-Neptunes
start their lives with large primordial atmopsheres that are are lost
over the first gigayear
\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_core-powered_2018}.
The observation from starspot-induced TTVs that the planetary orbit
could be aligned with the stellar spin differs from a number of other
close-in
mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
Overall however, it is consistent with a quiescent disk-driven
migration history for the planet, while the in-situ formation scenario
is disfavored because the necessary disk mass would always drive
planets to migrate
\citep{inamdar_formation_2015,ogihara_reassessment_2015}.

@@

\facilities{
 	{\it Astrometry}:
 	Gaia \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2020_edr3}.
 	{\it Imaging}:
    Second Generation Digitized Sky Survey. %,
    %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
 	%Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	{\it Spectroscopy}:
	CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
  %PFS ({\bf CITE}),
  %  MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
	AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
 	%Keck:I~(HIRES; \citealt{vogt_hires_1994}).
 	VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
%	  ASTEP:0.40$\,$m (ASTEP400),
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
%	  El Sauce:0.356$\,$m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
% 	NTT,
% 	SOAR (SOI),
 	TESS \citep{ricker_transiting_2015}.
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
}

>>>>

\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{---}
\revised{---}
\accepted{---}
\submitjournal{AAS Journals}
\shorttitle{Kepler\,1627}

@@

% \begin{abstract}
%   Kepler\,1627A is a G8V star previously found to host a
%   $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
%   selected for observation by the Kepler satellite because it is
%   nearby ($d\approx 333\,{\rm pc}$) and it resembles the Sun.  Combining
%   data from Gaia with rotation periods from TESS, we show that
%   Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
%   cluster.  Kepler\,1627Ab is therefore the youngest planet with a
%   precise age found to date by the main Kepler mission.
%   Analyzing the photometry, we find evidence that starspot crossings
%   distort the average transit shape, and induce transit timing
%   variations that suggest the planet is on a prograde orbit.  The 
%   flare arrival time distribution is most likely Poissonian, though
%   there may be hints ($<3$-$\sigma$) of additional flares at the
%   orbital and synodic periods.  In the future, spectroscopic studies
%   of \pn\ may help clarify the evolution of this mini-Neptune's orbit
%   and atmosphere.  More broadly, the combination of Gaia and TESS data
%   offers an opportunity to expand the census of age-dated planets; we
%   provide a literature concatenation of young stars to support this
%   effort.
% \end{abstract}

% 195 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby ($d\approx
  333\,{\rm pc}$) and it resembles the Sun.  Combining Gaia kinematics
  with rotation periods from TESS, we show that Kepler\,1627 is a
  member of the $35\pm XX$\,Myr old $\delta$~Lyr cluster.
  Kepler\,1627Ab is therefore the youngest planet with a precise age
  yet found by the main Kepler mission.  The precision of the Kepler
  photometry lets us detect two anomalies.  First, unresolved
  starspot-crossing events distort the average transit shape, and
  induce transit timing variations that imply the planet is likely on
  a prograde orbit.  Second, although the flare arrival time
  distribution is probably Poissonian, there may be an excess of
  flares separated by the orbital and synodic periods, and linear
  combinations thereof.  The $\delta$~Lyr cluster is one of many
  stellar groups whose properties have been clarified using the Gaia
  data; many of the other new clusters also overlap with confirmed and
  candidate exoplanets.  Further studies that combine Gaia kinematics
  and TESS rotation periods are therefore well-poised to expand the
  census of age-dated planets, and our understanding of planetary
  evolution.
\end{abstract}

% While thousands of exoplanets have been discovered orbiting nearby
% stars, the vast majority of them are several billion years old.  This
% makes it difficult to test theories for the origins of the different
% families of planets, since many evolutionary processes are expected to
% operate on timescales of less than 100 million years.  The K2 and TESS
% missions have enabled advances on this front [1,2,3,4,5,6].  The
% Kepler mission however has not yielded any planets with precise ages
% below one gigayear [7,8].  Here, we combine kinematic data from Gaia
% with stellar rotation periods from TESS to show that Kepler 1627 is a
% member of the 35 ± 10 million year old δ Lyr cluster.  A
% $3.6\,R_\oplus$ mini-Neptune transits the G8V primary on a 7.2 day
% orbit, and it is therefore the youngest planet yet found by the main
% Kepler mission.  The precision of the Kepler photometry lets us detect
% two anomalies.  First, unresolved starspot-crossing events distort the
% average transit shape, and induce transit timing variations that imply
% the planet is likely on a prograde orbit [9].  Second, although the
% flare arrival time distribution is most likely Poissonian, there may
% be hints of an excess of flares at the orbital and synodic periods.
% The δ Lyr cluster is one of many stellar groups whose nature has been
% clarified using the Gaia data; many other new clusters are expected to
% overlap datasets of confirmed and candidate planets.  Further studies
% that combine Gaia kinematics and TESS rotation periods are therefore
% well-poised to expand the census of age-dated planets, and our
% understanding of planetary evolution.

\keywords{
  exoplanet evolution (491),
  stellar associations (1582),
  open star clusters (1160),
	stellar ages (1581),
}

@@

\section{Introduction}

While thousands of exoplanets have been discovered orbiting nearby
stars, the vast majority of them are several billion years old.  This
makes it difficult to test origin theories for the different
families of planets, since many evolutionary processes are expected to
operate on timescales of less than 100 million years.

%TODO: better citations on radius evolution?
For instance, the ``mini-Neptunes'', thought to be made of
molten rocky cores \citep{kite_atmosphere_2020} and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years
\citep{owen_constraining_2020}.  Specifically,
they start with sizes of 4--12\,$R_\oplus$ at the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which can eventually deplete or entirely strip the envelope
from the rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
Discovering young planets and measuring their compositions and
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of mature
exoplanets \citep{Fulton_et_al_2017}.

The K2 and TESS missions have now enabled the detection of about ten
close-in planets younger than 100 million years, all smaller than
Jupiter
\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020}.
The Kepler mission however has not yielded any planets with precise
ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
at the time of the main Kepler mission (2009--2013), only four open
clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Since that time, analyses of the
kinematic, photometric, and astrometric Gaia data have expanded our
knowledge of open cluster and moving group memberships \citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,kerr_stars_2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis by J.
Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
the cluster from the Kepler-52 and Kepler-968 systems
\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
discovery, and the focus of this work, is that Kepler observed a
portion of the $\delta$\,Lyr cluster \citep[Stephenson-1;][]{stephenson_possible_1959,eggen_photometric_1968}.
The overlap between the Kepler field and the cluster is shown in
Figure~\ref{fig:skychart}.

A previous clustering analysis of the Gaia data
reported that
Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was a member of the
\cn\ \citep{kounkel_untangling_2019}.
Given the statistical validation of the close-in mini-Neptune
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018}, we therefore began by examining the age of
the star and its putative host cluster (Section~\ref{sec:cluster}). 
We find that the $\delta$ Lyr cluster is $\approx35$\,Myr old, and
that \sn\ is indeed a member of the cluster.
Focusing on the planet itself (Section~\ref{sec:planet}), we confirm
that despite a previously unreported binary companion, it is indeed a
mini-Neptune orbiting the roughly Sun-like primary star.
The presence of a correlation between transit timing variations and
the local light curve slope renders any other explanation implausible.
The broader implications for our ability to age-date a statistical
sample of planets are highlighted in closing (Section~\ref{sec:conc}).

\section{The Cluster}
\label{sec:cluster}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
    points are kinematically selected members of the $\delta$\,Lyr
    cluster.  Both Kepler (gray squares) and TESS (colored points)
    observed portions of the cluster.  Naked-eye stars ($m_{\rm
    V}<6.5$) are shown in gray; four of them (orange points) have
    their names annotated.  Kepler\,1627 (green star) was observed
    during the entirety of the Kepler mission, and has been observed
    by TESS for two lunar months to date.
    {\bf todo: move colorbar...}
		\label{fig:skychart}
	}
\end{figure}

@@

{\bf Jason Curtis todo}
We collected the previously measured
{\bf XXX} Kepler rotation periods (CITEP MCQUILLAN2014), and measured
{\bf YYY} new stellar rotation periods for cluster members observed by
TESS.  We also performed an isochronal analysis.  The results are
shown in Figure~\ref{fig:age}.  From the ensemble perspective, which
provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
of the star is $XX \pm YY$\,Myr.  Individual measurements of the
star's rotation period and lithium abundance corroborate this
conclusion.

\section{The Planet}
\label{sec:planet}

@@

\begin{figure}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f11.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Evidence for a prograde orbit from transit timing variations.}
    The time of each transit was measured, along with the local slope
    of the light curve.  The two quantities appear anti-correlated,
    which is most easily explained by starspot crossings during the
    first (second) half of transit inducing a positive (negative) TTV,
    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
    units along the abscissa are most easily understood by considering
    that the flux from the star changes by $\sim$60\,ppt per half
    rotation period ($\sim$1.3\,days).
		\label{fig:ttvslope}
	}
\end{figure}

@@

\section{Discussion \& Conclusions}
\label{sec:conc}

@@

In the context of the known transiting planets, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precisely measured ages spanning 10 to 100\,Myr include K2-33b,
DS~Tuc~Ab, HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet
V1298~Tau system
\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020}.
Of these systems, \pn\ inhabits a similar range of planetary size and
orbital period.  It is the smallest though ($3.6\,R_\oplus$), which
may hint at selection effects in the sample of young planets
\citep{zhou_2021_tois}.  Assuming that these planets have masses
between $\approx$10 and 20$\,M_\oplus$, their large sizes and presumed
low densities are in accord with the expectation that mini-Neptunes
start their lives with large primordial atmopsheres that are are lost
over the first gigayear
\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.
The observation from starspot-induced TTVs that the planetary orbit
could be aligned with the stellar spin differs from a number of other
close-in
mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
Overall however, it is consistent with a quiescent disk-driven
migration history for the planet, while the in-situ formation scenario
is disfavored because the necessary disk mass would always drive
planets to migrate
\citep{inamdar_formation_2015,ogihara_reassessment_2015}.

@@

\facilities{
 	{\it Astrometry}:
 	Gaia \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2020_edr3}.
 	{\it Imaging}:
    Second Generation Digitized Sky Survey. %,
    %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
 	%Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	{\it Spectroscopy}:
	CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
  %PFS ({\bf CITE}),
  %  MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
	AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
 	%Keck:I~(HIRES; \citealt{vogt_hires_1994}).
 	VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
%	  ASTEP:0.40$\,$m (ASTEP400),
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
%	  El Sauce:0.356$\,$m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
	  Kepler \citep{borucki_kepler_2010},
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
% 	NTT,
% 	SOAR (SOI),
 	TESS \citep{ricker_transiting_2015}.
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
}
